\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-02-22 intro.tex
removed references to probabilistic state estimation as neither t-rex or europa are formally handling that.  We may have demonstrated that a HMM can be modeled and executed within nddl but still neither of the 2 frameworks where aware of it.

At the Monterey Bay Aquarium Research Institute (MBARI), a unique
inter-disciplinary collaboration between biologists, ecologists,
geneticists and roboticists on the \can (Controlled Agile Novel
Observation Network) \cite{canon} has driven autonomous system
exploration and control. Fundamental questions related to sampling
strategies have driven the use cases for advanced concepts in
autonomy, specifically to Artificial Intelligence (AI) Planning and
Execution. Current work is also looking at how other fields in Machine
Intelligence especially Machine Learning can be leveraged to impact
the science of engineering systems in marine robotics and as a
consequence oceanography itself.

@@

AUV control architectures have primarily been driven by reactive
Subsumption-based approaches \cite{brooks86}. They have been adequate
to doing routine straight-line survey tracts which have played a
valuable role in the acceptance of AUVs as a new tool for ocean
science and DoD applications. However, scaling of problem domains in
addition to dealing with event-response and discovery in the
water-column challenges \emph{how} controllers are written in the
first place, maintained and upgraded over the life-cycle of the
platform. Dealing with off-nominal endogenous or exogenous conditions
including sensor failures stretches how AUVs are currently being
utilized; one consequence for instance, is any off-nominal condition
results in a \emph{fail-safe} surface mode to radio for help. Further
these approaches, words to the contrary are not only hard to maintain,
are not 'adaptive' enough. Despite vigorous defense by Brooks early on
\cite{Brooks91intelligencewithoutrea,Brooks91intelligencewithoutrep}
require a model if not of the world then of the \emph{physics} of the
vehicle when it interacts with the world.

Persistent and early legitimate criticism of deliberation on the other
hand also well articulated by Brooks
\cite{Brooks91intelligencewithoutrea,Brooks91intelligencewithoutrep}
was targeting the performance of in-situ deliberation. Terrestrial
robotic platforms were well-known to frequently stop and ``plan''. In
the marine robotics world, such a situation would simply not be
permissible given the need to make continuous observations without
stopping an under-articulated robotic platform such as an AUV. However
algorithmic and implementational advances by the AI community over the
years has produced a range of planners with demonstrated embedded
capabilities
\cite{simmons94,Haigh98,alami:1998p820,chien00,mus98,teichteil07}. Our
work has been influenced by and in turn influenced a wide range of
activities in the field of AI Planning and Plan Execution with
demonstrated real-world and highly visible capabilities in the Space
domain \cite{mus98,rajan00,aichang04,bresina05} for NASA. Lessons
learned have been applied to the world of marine robotics which we
bring together in this chapter. Important lessons that were derived
from this rich legacy involved the following:

@@

We have developed, tested and deployed the Teleo-Reactive EXecutive
(\texttt{T-REX}) an on-board adaptive control system that integrates
AI based planning and probabilistic state estimation in a hybrid
executive \cite{mcgann08a,mcgann08b,py10}.  Probabilistic state
estimation integrates a number of science observations to produce a
likelihood that the vehicle sensors perceive a feature of
interest. Onboard planning and execution enables adaptation of
navigation and instrument control based on the probability of having
detected such a phenomenon. It further enables goal-directed
commanding within the context of projected mission state and allows
for replanning for off-nominal situations and opportunistic science
events. The framework in addition to being used on an AUV, is general
enough be used for controlling a personal robot \cite{mcgann2009} and
is deployed on a European planetary rover testbed
\cite{goac11}. Probabilistic state estimation is discussed elsewhere
\cite{mcgann08d} and out of scope for this chapter. Deliberation and
reaction are integrated systematically over different temporal and
functional scopes within a single agent and a single model that covers
the needs of high-level mission management, low-level navigation,
instrument control, and detection of unstructured and poorly
understood phenomena. \rx is deployed on MBARI's Dorado AUV shown in
Fig. \ref{fig:auv-fig}, which to the best of our knowledge is the only
operational marine platform anywhere being used for routine scientific
surveys with onboard plan synthesis.

>>>>

At the Monterey Bay Aquarium Research Institute (MBARI), a unique
inter-disciplinary collaboration between biologists, ecologists,
geneticists and roboticists on the \can (Controlled Agile Novel
Observation Network) program \cite{canon} has driven autonomous system
exploration and control. Fundamental questions related to sampling
strategies have driven the use cases for advanced concepts in
autonomy, specifically to Artificial Intelligence (AI) Planning and
Execution. Current work is also looking at how other fields in Machine
Intelligence especially Machine Learning can be leveraged to impact
the science of engineering systems in marine robotics and as a
consequence oceanography itself.

@@

AUV control architectures have primarily been driven by reactive
Subsumption-based approaches \cite{brooks86}. They have been adequate
to doing routine straight-line survey tracts which have played a
valuable role in the acceptance of AUVs as a new tool for ocean
science and DoD applications. However, scaling of problem domains in
addition to dealing with event-response and discovery in the
water-column challenges \emph{how} controllers are written in the
first place, maintained and upgraded over the life-cycle of the
platform. Dealing with off-nominal endogenous or exogenous conditions
including sensor failures stretches how AUVs are currently being
utilized; one consequence for instance, is any off-nominal condition
results in a \emph{fail-safe} surface mode to radio for help. Further
these approaches, words to the contrary are not only hard to maintain,
are not 'adaptive' enough. Despite vigorous defense by Brooks early on
\cite{Brooks91intelligencewithoutrea,Brooks91intelligencewithoutrep},
autonomous systems require a model if not of the world then of the 
\emph{physics} of the vehicle when it interacts with the world.

Persistent and early legitimate criticism of deliberation on the other
hand also well articulated by Brooks
\cite{Brooks91intelligencewithoutrea,Brooks91intelligencewithoutrep}
was targeting the performance of in-situ deliberation. Terrestrial
robotic platforms were well-known to frequently stop and ``plan''; 
assuming in the meantime that the world will remain static. In
the marine robotics world, such a situation would simply not be
permissible given the need to make continuous observations without
stopping an under-articulated robotic platform such as an AUV. However
algorithmic and implementational advances by the AI community over the
years has produced a range of planners with demonstrated embedded
capabilities
\cite{simmons94,Haigh98,alami:1998p820,chien00,mus98,teichteil07}. Our
work has been influenced by and in turn influenced a wide range of
activities in the field of AI Planning and Plan Execution with
demonstrated real-world and highly visible capabilities in the Space
domain \cite{mus98,rajan00,aichang04,bresina05} for NASA. Lessons
learned have been applied to the world of marine robotics which we
bring together in this chapter. Important lessons that were derived
from this rich legacy involved the following:

@@

We have developed, tested and deployed the Teleo-Reactive EXecutive
(\texttt{T-REX}) an on-board adaptive control system that integrates
AI based planning and %\sout{probabilistic} 
state estimation in a hybrid executive\cite{mcgann08a,mcgann08b,py10}. 
%  \sout{Probabilistic state estimation integrates a number of science observations to produce a
% likelihood that the vehicle sensors perceive a feature of
% interest}.\fcomment{\rx {\em does not} integrate probabilistic state
%   estimation! ... The HMM estimation was coded on top of \eu model but
%   neither \eu or \rx had explicit management of it} 
State estimation allow the system to keep track of the world evolution
as time advances based on a formal model distributed through the architecture.
Onboard planning and execution enables adaptation of
navigation and instrument control based on the state identified during
state estimation.
% \sout{the probability of having detected such a phenomenon} 
It further enables goal-directed
commanding within the context of projected mission state and allows
for replanning for off-nominal situations and opportunistic science
events. The framework in addition to being used on an AUV, is general
enough be used for controlling a personal robot \cite{mcgann2009} and
is deployed on a European planetary rover testbed
\cite{goac11}. While it is possible to integrate probabilistic
estimation on top of this framework \cite{mcgann08d}, such aspect is
out of scope of this chapter.
% Probabilistic state estimation through modeling within 
% this framework can be done is discussed elsewhere
% \cite{mcgann08d} and out of scope for this chapter. 
Deliberation and reaction are integrated systematically over different temporal and
functional scopes within a single agent and a single model that covers
the needs of high-level mission management, low-level navigation,
instrument control, and detection of unstructured and poorly
understood phenomena. \rx is deployed on MBARI's Dorado AUV shown in
Fig. \ref{fig:auv-fig}, which to the best of our knowledge is the only
operational marine platform anywhere being used for routine scientific
surveys with onboard plan synthesis.
